Infinite diameter confidence sets in the selection function model for publication bias

### 2020-04-24 infiniteci.tex
:pear: Some small changes.

\title{Infinite diameter confidence sets in a model for publication bias}

@@

\titletag{An arXiv preprint, v1.1}
\makeatletter
\titlerunning{\@title}
\makeatother
\authorrunning{Jonas Moss}

\begin{document}
\maketitle
\begin{abstract}
Meta-analysis is the quantitative combination of results from different studies. The assumptions of classical meta-analysis models are not satisfied whenever publication bias is present, which causes inconsistent parameter estimates. A popular method to account for publication bias in meta-analysis is the selection function model, but it has been noted that frequentist estimation of this model is problematic. This paper shows there is no confidence set of guaranteed finite diameter for the mean and heterogeneity parameters in the selection function model where the selection function is a step function. The proof is based on a generalization of the Gleser--Hwang theorem.
\end{abstract}

@@

Meta-analysis is the quantitative combination of results from different studies \citep{lipsey2001practical} with the goal of reducing sampling error.
Publication bias occurs when the published research literature is
not representative of the all the research that has actually been
done \citep{rothstein2006publication}. An important case of publication
bias is when publication decisions are made based on \textit{p}-values
\citep{sterling1959publication}, which (and its sister phenomenon, \textit{p}-hacking \citep{simmons2011false})
can cause seriously biased conclusions when not accounted
for \citep{simmons2011false,moss2019modelling}.

One of the most important models for meta-analysis is the normal random effects
model with normal likelihoods where the
standard deviations are assumed to be known \citep{hedges1998fixed}.
The likelihood for an observation in this model is $\phi(x_{i}\mid\theta_{0},(\sigma_{i}^{2}+\tau^{2})^{1/2})$,
where $\phi(x\mid\mu,\sigma)$ is the normal density with
mean $\mu$ and standard deviation $\sigma$. The parameter $\tau$ is the standard deviation of the random effects distribution, and is known as the heterogeneity parameter, while $\sigma_i$ are the study specific standard errors and $\theta_0$ is the mean of the effect size distribution. A modification of the random
effects meta-analysis model essentially due to \citet{hedges1984estimation}
allows us to account for \textit{p}-value based publication bias.

@@

Frequentist estimation of the step function publication bias model
is problematic, as noted by for instance \citet[appendix, 1]{mcshane2016adjusting}.
The purpose of this paper is to formalize and prove exactly how bad frequentist estimation of this model can be. I do this by proving there are no confidence sets of guaranteed finite diameter for the mean parameter $\theta_{0}$
and the heterogeneity parameter $\tau$ for any coverage $1-\alpha$.
This is a problematic result for two reasons: (i) It would be hopeless
to report confidence sets for $\tau^{2}$ like $[0.5,\infty)$,
as they have no practical value. (2) It shows that the automatic confidence
sets procedures that are guaranteed to yield finite confidence sets
of some positive nominal coverage, such as bootstrapped confidence
sets, likelihood-ratio based confidence sets, and subsampling confidence
sets never have true coverage greater than $0$ \citep[see][]{gleser996bootstrap}.

@@

\section{Definitions and Proofs}

Let $P_{\theta},\theta\in\Theta$ be a family of dominated probability
measures with densities $p_{\theta}$. Let $g:\Theta\to E$ for some
set $E$ equipped with a positive function $\left\Vert \cdot\right\Vert :E\to[0,\infty)$.
Here $g$ maps $\theta$ to the parameter we wish to form a confidence
set for, sometimes called a \textit{focus parameter} ((cite Nils)). For instance, $g(\theta) = \theta$ or $g(\theta) = E_\theta(X)$, the expectation of some random variable $X$ calculated under $P_\theta$.

@@

The problem with non-existence
of confidence sets with guaranteed finite diameter for finite-dimensional
parameters has not received much attention in the statistical literature,
as lamented by \citet{gleser996bootstrap}. Some references in this
area are \citet{bahadur1956nonexistence}, who studied infinitely
large confidence sets in non-parametric estimation of the mean, \citeauthor{romano2004non}'s
(2004) extension of their results, \citet{Donoho1988-hg} and \citet{Pfanzagl1998-fe}.
\begin{lem}
Let $C$ be a confidence set and $A_{g(\theta)}$ its associated
acceptance sets. If there exists a sequence $\left\{ \theta_{n}\right\} _{n=1}^{\infty}$
with $\left\Vert g(\theta_{n})\right\Vert \to\infty$ such
that $\lim_{n\to\infty}P_{\theta}(A_{g(\theta_{n})})>0,$then
the diameter of $C$ at $\theta$ is infinite with positive probability.
\end{lem}
\begin{proof}
Assume without loss of generality that $\left\Vert g(\theta_{n})\right\Vert \geq n$
for each $n$; such a sequence can always be found by appropriately
filtering the original $\left\{ \theta_{n}\right\} _{n=1}^{\infty}$.
Since $\left\{ D\geq n\right\} $ is a decreasing sequence of sets,
$P_{\theta}(D=\infty)=\lim_{n\to\infty}P_{\theta}(D\geq n)$
for each $\theta$. From $A_{g(\theta_{n})}\subseteq\left\{ D\geq n\right\} $
it follows that $P_{\theta}(A_{g(\theta_{n})})\leq P_{\theta}(D\geq n)$,
thus $P_{\theta}(D=\infty)=\lim_{n\to\infty}P_{\theta}(D\geq n)\geq\lim_{n\to\infty}P_{\theta}(A_{g(\theta_{n})})>0$.
\end{proof}
This is the extension of the Gleser--Hwang theorem.
\begin{thm}
\label{thm:Infinite diameter main theorem}If there exists a sequence
$\left\{ \theta_{i}\right\} _{i\in\mathbb{N}}$ with $\left\Vert g(\theta_{i})\right\Vert \to\infty$
such that $p_{\theta_{i}}\to f^{\star}$ pointwise and $\textrm{supp}p_{\theta}=\textrm{supp}f^{\star}$
for each $\theta$, then every confidence set with coverage $1-\alpha>0$
has infinite diameter with positive probability.
\end{thm}
\begin{proof}
Assume without loss of generality that $\left\Vert g(\theta_{n})\right\Vert \geq n$.
By a variant of the dominated convergence theorem \citep[Exercise 16.4a]{billingsley1995probability}, $\int_{D\geq n}p_{\theta_{n}}d\mu\to\int_{D=\infty}f^{\star}d\mu$.
Since $A_{g(\theta_{n})}\subseteq\left\{ D\leq n\right\}$, we have 
$1-\alpha\leq\int_{A_{g(\theta_{n})}}p_{\theta_{n}}d\mu\leq\int_{D\geq n}p_{\theta_{n}}d\mu$
for all $n$, thus $\int_{D=\infty}f^{\star}d\mu>1-\alpha>0$. The
result follows from $P_{\theta}(D\geq n)=\int_{D\geq n}(p_{\theta}/p_{\theta_{n}})p_{\theta_{n}}d\mu>\int_{D=\infty}(p_{\theta}/f^{\star})f^{\star}d\mu>0$.
\end{proof}
This an extension of Theorem \ref{thm:Infinite diameter main theorem}
to mixture distributions.
\begin{prop}
\label{cor:Mixture model corollary}Let $f_{\theta}^{1},f_{\theta}^{2},\ldots$
be a finite or infinite sequence of densities, $\pi(\theta)$
a countable probability vector and $\sum_{n=1}^{\infty}\pi_{n}(\theta)f_{\theta}^{n}$
a mixture distribution. Assume there is a sequence $\left\{ \theta_{i}\right\} _{i\in\mathbb{N}}$
with $\left\Vert g(\theta_{i})\right\Vert \to\infty$ such
that $f_{\theta_{i}}^{1}\to f^{\star}$ pointwise and $\textrm{supp}f_{\theta}^{1} = \textrm{supp}f^{\star}$
for each $\theta$ . If there is a parameter $\theta'\in\Theta$ satisfying
$\pi_{1}(\theta')>\alpha$, the mixture $\sum_{n=1}^{\infty}\pi_{n}(\theta)f_{\theta}^{n}$
admits no almost surely finite diameter confidence set of coverage
$1-\alpha$ for $g(\theta)$.
\end{prop}

>>>>

\title{Infinite diameter confidence sets in the selection function model for publication bias}

@@

\titletag{An arXiv preprint, v2}
\makeatletter
\titlerunning{\@title}
\makeatother
\authorrunning{Jonas Moss}

\begin{document}
\maketitle
\begin{abstract}
Meta-analysis is the quantitative combination of results from different studies. The assumptions of classical meta-analysis models are not satisfied whenever publication bias is present, which causes inconsistent parameter estimates. A popular method to account for publication bias in meta-analysis is the selection function model, but frequentist estimation of this model is problematic. This paper shows there is no confidence set of guaranteed finite diameter for the mean and heterogeneity parameters in the selection function model. The proof is based on a generalization of the Gleser--Hwang theorem.
\end{abstract}

\keywords{meta-analysis \and publication bias \and selection function model \and p-hacking}

@@

Meta-analysis is the quantitative combination of results from different studies with the goal of reducing sampling error \citep{lipsey2001practical}.
Publication bias occurs when the published research literature is
not representative of the all the research that has actually been
done \citep{rothstein2006publication}. An important case of publication
bias is when publication decisions are made based on \textit{p}-values
\citep{sterling1959publication}, which, along its sister phenomenon \textit{p}-hacking \citep{simmons2011false},
can cause seriously biased conclusions when not accounted
for \citep{simmons2011false,moss2019modelling}.

One of the most important models for meta-analysis is the normal random effects
model with normal likelihoods where the
standard deviations are assumed to be known \citep{hedges1998fixed}.
The likelihood for an observation in this model is $\phi(x_{i}\mid\theta_{0},(\sigma_{i}^{2}+\tau^{2})^{1/2})$,
where $\phi(x\mid\mu,\sigma)$ is the normal density with
mean $\mu$ and standard deviation $\sigma$. The parameter $\tau$ is the standard deviation of the random effects distribution, known as the heterogeneity parameter. The $\sigma_i$s are the study-specific standard errors, and $\theta_0$ is the mean of the effect size distribution. A modification of the random effects meta-analysis model due to \citet{hedges1984estimation}
allows us to account for \textit{p}-value based publication bias.

@@

Frequentist estimation of the step function publication bias model
is problematic, as noted by \citet[Appendix, 1]{mcshane2016adjusting}.
The purpose of this paper is to formalize and prove exactly how bad frequentist estimation of this model can be. I do this by proving there are no confidence sets of guaranteed finite diameter for the mean parameter $\theta_{0}$
and the heterogeneity parameter $\tau$ for any coverage $1-\alpha$.
This is a problematic result for two reasons: (i) It would be hopeless
to report confidence sets for $\tau^{2}$ like $[0.5,\infty)$,
as they have no practical value. (ii) It shows that the automatic confidence
sets procedures that are guaranteed to yield finite confidence sets
of some positive nominal coverage, such as bootstrapped confidence
sets, likelihood-ratio based confidence sets, and subsampling confidence
sets never have true coverage greater than $0$ \citep[see][]{gleser996bootstrap}.

@@

\section{Definitions and Proofs}
Let $P_{\theta},\theta\in\Theta$ be a family of dominated probability
measures with densities $p_{\theta}$. Let $g:\Theta\to E$ for some
set $E$ equipped with a positive function $\left\Vert \cdot\right\Vert :E\to[0,\infty)$.
Here $g$ maps $\theta$ to the parameter we wish to form a confidence
set for, sometimes called a \textit{focus parameter} ((cite Nils)). For instance, $g(\theta) = \theta$ or $g(\theta) = E_\theta(X)$, the expectation of some random variable $X$ calculated under $P_\theta$.

@@

The problem with non-existence
of confidence sets with guaranteed finite diameter for finite-dimensional
parameters has not received much attention in the statistical literature,
as lamented by \citet{gleser996bootstrap}. Some references in this
area are \citet{bahadur1956nonexistence}, who studied infinitely
large confidence sets in non-parametric estimation of the mean, \citeauthor{romano2004non}'s
(2004) extension of their results, \citet{Donoho1988-hg} and \citet{Pfanzagl1998-fe}.
\begin{lem}
Let $C$ be a confidence set and $A_{g(\theta)}$ its associated
acceptance sets. If there exists a sequence $\left\{ \theta_{n}\right\} _{n=1}^{\infty}$
with $\left\Vert g(\theta_{n})\right\Vert \to\infty$ such
that $\lim_{n\to\infty}P_{\theta}(A_{g(\theta_{n})})>0,$then
the diameter of $C$ at $\theta$ is infinite with positive probability.
\end{lem}
\begin{proof}
Assume without loss of generality that $\left\Vert g(\theta_{n})\right\Vert \geq n$
for each $n$; such a sequence can always be found by appropriately
filtering the original $\left\{ \theta_{n}\right\} _{n=1}^{\infty}$.
Since $\left\{ D\geq n\right\} $ is a decreasing sequence of sets,
$P_{\theta}(D=\infty)=\lim_{n\to\infty}P_{\theta}(D\geq n)$
for each $\theta$. From $A_{g(\theta_{n})}\subseteq\left\{ D\geq n\right\} $
it follows that $P_{\theta}(A_{g(\theta_{n})})\leq P_{\theta}(D\geq n)$,
thus $P_{\theta}(D=\infty)=\lim_{n\to\infty}P_{\theta}(D\geq n)\geq\lim_{n\to\infty}P_{\theta}(A_{g(\theta_{n})})>0$.
\end{proof}
This is the extension of the Gleser--Hwang theorem.
\begin{thm}
\label{thm:Infinite diameter main theorem}If there exists a sequence
$\left\{ \theta_{i}\right\} _{i\in\mathbb{N}}$ with $\left\Vert g(\theta_{i})\right\Vert \to\infty$
such that $p_{\theta_{i}}\to f^{\star}$ pointwise and $\textrm{supp}p_{\theta}=\textrm{supp}f^{\star}$
for each $\theta$, then every confidence set with coverage $1-\alpha>0$
has infinite diameter with positive probability.
\end{thm}
\begin{proof}
Assume without loss of generality that $\left\Vert g(\theta_{n})\right\Vert \geq n$.
By a variant of the dominated convergence theorem \citep[Exercise 16.4a]{billingsley1995probability}, $\int_{D\geq n}p_{\theta_{n}}d\mu\to\int_{D=\infty}f^{\star}d\mu$.
Since $A_{g(\theta_{n})}\subseteq\left\{ D\leq n\right\}$, we have 
$1-\alpha\leq\int_{A_{g(\theta_{n})}}p_{\theta_{n}}d\mu\leq\int_{D\geq n}p_{\theta_{n}}d\mu$
for all $n$, thus $\int_{D=\infty}f^{\star}d\mu>1-\alpha>0$. The
result follows from $P_{\theta}(D\geq n)=\int_{D\geq n}(p_{\theta}/p_{\theta_{n}})p_{\theta_{n}}d\mu>\int_{D=\infty}(p_{\theta}/f^{\star})f^{\star}d\mu>0$.
\end{proof}
This is an extension of Theorem \ref{thm:Infinite diameter main theorem}
to mixture distributions.
\begin{prop}
\label{cor:Mixture model corollary}Let $f_{\theta}^{1},f_{\theta}^{2},\ldots$
be a finite or infinite sequence of densities, $\pi(\theta)$
a countable probability vector and $\sum_{n=1}^{\infty}\pi_{n}(\theta)f_{\theta}^{n}$
a mixture distribution. Assume there is a sequence $\left\{ \theta_{i}\right\} _{i\in\mathbb{N}}$
with $\left\Vert g(\theta_{i})\right\Vert \to\infty$ such
that $f_{\theta_{i}}^{1}\to f^{\star}$ pointwise and $\textrm{supp}f_{\theta}^{1} = \textrm{supp}f^{\star}$
for each $\theta$ . If there is a parameter $\theta'\in\Theta$ satisfying
$\pi_{1}(\theta')>\alpha$, the mixture $\sum_{n=1}^{\infty}\pi_{n}(\theta)f_{\theta}^{n}$
admits no almost surely finite diameter confidence set of coverage
$1-\alpha$ for $g(\theta)$.
\end{prop}
